---
title: "Claw analysis"
output: html_notebook
author: Andrew L Jackson
---



Attempt to get to the bottom of the data patterns and Kerry's concerns.

## Setup

```{r}
library(SIBER)
library(tidyverse)
```


## Import data

```{r}

data <- read_csv("SI 18MAR22 TipBaseMid.csv")
names(data)

claw_df <- as.data.frame(select(data, iso1, iso2, group, community))

```

## Plot the raw data
```{r}

g1 <- ggplot(data = claw_df, 
             mapping = aes(x = iso1, y = iso2, 
                           color = group,
                           shape = community)) + 
  geom_point() + 
  stat_ellipse(aes(linetype = community))

print(g1)

```

## Imoprt to SIBER and generate some basic stats

```{r}
siber_claw<-createSiberObject(claw_df)

siber_claw$sample.sizes



```


plot the siber object to visualise both the community and group data to make sense of the statistics

```{r}

plotSiberObject(siber_claw, 
                hulls = TRUE, 
                ellipses = TRUE)

```

From this is appears that the East community of points has a larger triangle joining the midpoints of its three ellipses compared with the West community which is so small its practically a line. This suggests that almost all the community metrics will be larger for East than West. The community metrics are based on the mid-points of each group within a community, not the entire set of points that make up that community: the reasons for this are articulated in the original paper. 

On the other hand, the ellipses for West are much larger than East and so the group level metrics will follow that pattern of West > East. 

## Thoughts

perhaps you intended the community metrics to be calculated on each of the ellipses? I would recommend this for reasons articulated int he original paper, but it _is_ possible to do it computationally


```{r}

# split the claw data into a list by both group and community.
list_claw <- claw_df %>% 
  split(interaction(.$group, .$community)) #%>% map(laymanMetrics(x = .$iso1, y = .$iso2))


# preallocate group_layman as an empty list and name it same as list_claw
group_layman <- vector("list", length(list_claw))
names(group_layman) <- names(list_claw)

# loop over all the group-community combiations and calculate the layman metrics
for (i in 1:length(list_claw)){
  
  tmp <- list_claw[[i]]
  group_layman[[i]] <- laymanMetrics(tmp$iso1, tmp$iso2)
  
  
}


print(group_layman)

```



