---
title: "Emmi Kurosawa's spring/fall data"
output: html_notebook
---


I am going to analyse these data with simmr instead of mixsiar for speed for model fitting.

## Setup

```{r}
library(simmr)
library(tidyverse)
```

## Import the data

```{r}

mix_fall      <- read_csv("mix_Fall.csv") %>% as.matrix()
mix_spring    <- read_csv("mix_Spring.csv") %>% as.matrix()

source_fall   <- read_csv("source_Fall.csv")
source_spring <- read_csv("source_Spring.csv")

tdfs          <- read_csv("tef.csv")

```

Extract data to simmr formatting requirements

```{r}
src_mean_fall <- source_fall %>% select(Meand13C, Meand15N) %>%
  as.matrix()
src_mean_spring <- source_spring %>% select(Meand13C, Meand15N) %>%
  as.matrix()

src_sd_fall <- source_fall %>% select(SDd13C, SDd15N) %>%
  as.matrix()
src_sd_spring <- source_spring %>% select(SDd13C, SDd15N) %>%
  as.matrix()

src_names_fall <- source_fall$Sources
src_names_spring <- source_spring$Sources

tdf_means <- tdfs %>% select(Meand13C, Meand15N) %>% as.matrix()
tdf_sds   <- tdfs %>% select(SDd13C, SDd15N) %>% as.matrix()

```


## Create the simmr objects

First we can do "Fall".

```{r}

simmr_in_fall <- simmr_load(
  mixtures = mix_fall,
  source_names     = src_names_fall,
  source_means     = src_mean_fall,
  source_sds       = src_sd_fall,
  correction_means = tdf_means,
  correction_sds   = tdf_sds
  )


plot(simmr_in_fall)
```
Second we can do "Spring".

```{r}


simmr_in_spring <- simmr_load(
  mixtures = mix_spring,
  source_names     = src_names_spring,
  source_means     = src_mean_spring,
  source_sds       = src_sd_spring,
  correction_means = tdf_means,
  correction_sds   = tdf_sds
  )


plot(simmr_in_spring)
```

## Fit the models

```{r}

custom_mcmc <- list(iter = 100000, 
                    burn = 1000, 
                    thin = 10, 
                    n.chain = 4)

simmr_out_fall <- simmr_mcmc(simmr_in_fall, 
                             mcmc_control = custom_mcmc)
simmr_out_spring <- simmr_mcmc(simmr_in_spring, 
                               mcmc_control = custom_mcmc)


```

## Check the outputs

check convergence which all seems fine except the estimated residual error terms which are quite high. But lets keep going for now, as an even longer run should fix this. 

```{r}
# Check convergence - values should all be close to 1
summary(simmr_out_fall, type = "diagnostics")
summary(simmr_out_spring, type = "diagnostics")

```

Look at output


```{r}
summary(simmr_out_fall, type = "statistics")
summary(simmr_out_spring, type = "statistics")
```

Plot outputs
```{r}
plot(simmr_out_fall, type = "histogram")
plot(simmr_out_spring, type = "histogram")
```


There is good evidence of the posteriors diverging from the priors indicating the data is overriding the priors. Only Ref.Plant in Fall seems to be almost identical to its prior. 

```{r}
# Look at influence of priors
prior_viz(simmr_out_fall)
prior_viz(simmr_out_spring)

```
Check the correlations among the posterior proportions

```{r}
plot(simmr_out_fall, type = "matrix")
plot(simmr_out_spring, type = "matrix")

```







